Functional diversity improves tropical forest recovery from disturbance shifting from complementarity to selection effect (to rework)

### 2018-07-13 draft.Rmd
Draft 0.1.9005 Supplementary: RNE & LM

---
title: Functional diversity improves tropical forest resilience to disturbance
author: "Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabien Fischer, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  github_document: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
  bookdown::word_document2: default
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(tidyverse)
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 8,
                cache = T, cache.lazy = F)
```

@@

Natural disturbances caused by the death of one or more trees, are the dominant forms of forest renewal in tropical forests as the creation of canopy gaps continuously modifies forest horizontal and vertical structure [@Veblen1992]. Today, climate change and other direct human activities are modifying forest disturbance regimes, with huge and under evaluated impacts. When human induced disturbances are high, it is referred to as "degraded forests" characterized by a modification of the forest ecosystem functioning, possibly reducing the furnished ecosystem services, while keeping the ecosystem as a forest, contrary to deforestation [@Simula2009]. Besides deforestation has been rightly the focus of worldwide attention, degradation has been less studied and quantified through tropics. The main source of degradation around the tropics is forest logging that encompass selective timber harvest, in native forests, and wood fuel harvest, often in secondary forests [@Pearson2017]. Tropical forests disturbances account for 2.07 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005], estimating to cover areas 10 or more times larger than deforestation itself [@Herold2011]. Finally, disturbance in tropical forest dynamics induced by climate change likely results from significant changes in precipitation regime [@Davidson2012] with intense drought events resulting in huge mortality rates and biomass loss [@Phillips2009].

Tropical forest ecosystem response to disturbance depends first on the disturbance intensity itself. For instance, @DeAvila2015, focusing on timber and wood fuel harvest in Central Amazonia, found that when disturbance is high, e.g. induced by post logging sylvicultural practices had stronger negative effect on the post-disturbance system trajectory than logging itself. @Carreno-Rocabado2012 found no direct effect of disturbance intensity logging on community diversity itself but a community composition drift towards 'fast' acquisitive species [@reich_world-wide_2014]. However, even the short-term impact on biodiversity is unclear, as sometimes short term increase in diversity can occur at low disturbance intensities [@Martin2015].

The main reason of our lack of knowledge is the difficulty to conduct long term empirical study so that the use of forest simulators has become inevitable [@Huth2004]. Individual-based models of forest dynamics with species-explicit model parameters present the perfect framework to develop such joint biodiversity-ecosystem functioning studies [@Marechaux2017]. Starting from scratch, these models allow describing forest recovery through time, assessing individual tree recruitment, growth and mortality [@Bugmann2001]. Managing diversity in such models is still challenging and many models have built open Plant Functional Types to manage the volume of model parameters. [@Pacala1996].

Recently the forest growth simulator TROLL [@Chave1999], an individual-based and spatially-explicit forest model has been explicitly developed to deal with high diversity levels by using plant ecophysiological traits as surrogates of model parameters [@Marechaux2017]. Consequently, TROLL allow to fully simulate a highly-diverse system and offers the opportunity to study biodiversity-ecosystem functioning *in silico*.

Our work is based on the general hypothesis of a positive relationships between biodiversity and productivity. We assumed that when a disturbance event happen, due to a higher productivity, more diverse forests will recover quicker and thus be more resilient. This theoretical expectation stand on two processes: complementarity and selection effects [@Loreau2001]. Complementarity effect is the addition of better resources acquisition, due to resources partitioning and niche differentiation, and facilitation in diverse communities. Selection effect results from resources preemption by more competitive species. Moreover, we expect the sampling effect to improve resilience of species-rich forest, with redundancy and a lower risk to lose important functional traits.

Generally, positive relationship between biodiversity and productivity were empirically and experimentally demonstrated on grassland systems [@Hooper2005a; @Loreau2001a; @Naeem2002]; but few studies focused on the case of forests. One of the few studies was realized by @Chisholm2013 and has shown a positive significative relationship between species richness and wood productivity on a worldwide forest network. But those study still presents three major limits: (1) study time are inferior to a tree life time, (2) experimental network include scarcely disturbed plot, and (3) correlative approach does not explain mechanisms involved in the relationship.

In the present study, we thus focused on mechanisms involved in the relationship between biodiversity and forest ecosystem resilience. We used a simulation approach using TROLL model to assess long term processes for different levels of disturbances. We first assessed diversity effect on forest resilience of structure and functioning using numerous indices of taxonomic and functional diversities. Then, we measured biodiversity net effect, partitioned into complementarity and selection effects, on resilience of simulated tropical forest communities to disturbance.

@@

TROLL is an individual-based and spatially explicit forest growth model. TROLL simulates the life cycle of individual trees from recruitment, with a diameter at breast height (dbh) above 1 cm, to death with growth and seed production. Trees are growing in a spatialized light environment explicitly computed within voxels of 1 $m^3$. Each tree is consistently defined by its age, diameter at breast height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA). Tree geometry is calculated with allometric equations but leaf area vary dynamically within each crown following carbon allocations. Voxels resolution of 1 $m^3$ allow the establishment of maximum one tree by 1x1 m pixels. Each tree is flagged with a species label inherited from the parent tree through the seedling recruitment. A species label is associated to a number of species specific parameters (Table \@ref(tab:traits)) related to functional trait values. Carbon assimilation is computed over half-hourly period of a representative day. Then allocation is computed to simulate tree growth from an explicit carbon balance. Environment is updated at each time step set to one month. Seed-rains and seedlings are not simulated explicitly but as a pool. In addition below-ground processes, herbaceous plants, epiphytes and lianas are not simulated inside TROLL. The source code is written in C++ and available upon request. All modules of TROLL models are further detailed in @Marechaux2017. Following an underestimation of leaf lifespan for low LMA species, we designed our own leaf lifespan allometry, specific to tropical tree species, to be used in subsequent analysis (see supplementary materials S1).

@@

The disturbance module inside TROLL was designed in the simplest way by, for a given iteration $disturb_{iter}$, picking individual trees randomly up with a uniform law and without any individuals nor species targeting. Selected individuals are then removed without triggering a treefall to avoid any side effect. The operation is repeated until the disturbance result in a defined lost basal area ($disturb_{intensity}$).

## Design of experiment

In order to assess the role of biodiversity in ecosystem response to disturbance, we  simulated post-disturbance forest trajectories, encompassing both variation of disturbance intensity and biodiversity levels. The gradient of disturbance intensity was represented by percentage of basal area loss (0%, 25%, 50% and 75%). Biodiversity was assessed with two perspectives: taxonomic and functional. We used species richness $SR$ to simulate initial assemblages with diverse taxonomic diversity (5, 25, and 125 species). Functional diversity was assessed after running the initial simulations, *i.e.* before applying the disturbance. We focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$. For each level of species richness $SR$, we simulated 20 communities with a large range of convex hull volume $CHV$ but with a community weighted means close to the regional species pool community weighted means. This design of experiments resulted in 60 tree communities ($3~SR*20~CHV$) and 240 simulations ($60 ~communities*4~levels~of~disturbance$) over 600 years (maturity being assumed after 500 years of regeneration [@Marechaux2017]). Functional diversities at the end of simulations were assessed with @villeger_new_2008 indices (FRIC, FEve, FDiv, and FDis).

## Analysing Ecosystem Response

### Ecosystem structure and functioning

We described the simulated tropical forests from changes in values of forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and in values of forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$)).

The resilience of forest structure and functioning values was assessed with:

\begin{equation}
  R\left(t\right)=\frac{Recovery\left(t\right)}{Loss\left(t_d\right)} \approx \frac{X_T(t)}{X_C(t)}
  (\#eq:Resilience)
\end{equation}

The resilience of the system $R(t)$ at the time $t$ is described by the ratio of recovery $Recovery(t)$ at time $t$ to loss suffered $Loss(t_d)$ at disturbance time $t_d$. But in our peculiar case of simulated forest ecosystems, the equilibrium used to calculate $Loss(t_d)$ is dynamic in time so that we simulated an undisturbed control ecosystem $C$. And the resilience of the system $R(t)$ at the time $t$ was defined as the ratio of the ecosystem metric values in the disturbed simulation $X_T(t)$ over the ecosystem metric values from the control $X_C(t)$ (the control simulations being relatively stable due to small stochasticity among TROLL simulations [@Marechaux2017]). $R(t)$  will be equal to $R_{eq} = 1$ when reaching the equilibrium value. Next, we calculated an euclidean distance to equilibrium $d_{eq}(t)$ as $d_{eq}(t) = \sqrt{(R_{eq} - R(t))^2}$. Ecosystem euclidean distance to equilibrium was calculated in a multi-dimensional space for (the two functions described above: (i) forest structure (AGB, BA, N, N10, and N30) and (ii) forest functioning (GPP, NPP, Rday, and Rnight). We then integrate the euclidean distances over time to get our ecosystem resilience indicators.

We then applied the following to each indicator:

1. Normalizing the treatment values by their corresponding control (Figure \@ref(fig:datatrans) A and B).
2. Respectively gathering forest structure and forest functioning together to get synthetic indicators (Figure \@ref(fig:datatrans) C)
3. Integration over time (Figure \@ref(fig:datatrans) D)
4. The ranking of simulations regarding disturbance intensity was highly stable over time for the 240 simulations. So we used the cumulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem resilience.

```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation. Ecosystem outputs (**A**) are normalized by the control value over time to calculate resilience (**B**); resilience of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance at equilibrium is integrated over time in a cumulative sum (**D**). ', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
cols <- RColorBrewer::brewer.pal(3, "YlOrRd")
g_agb <- ecosystem %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb/1000, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Aboveground biomass (tonC/ha)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                       values=c("#999999", cols))
g_resilience <- resilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb, color = disturbance)) +
  geom_hline(yintercept = 1, col = 'red') +
  geom_point() +
  xlab('Time (years)') +
  ylab('Resilience of aboveground biomass') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_deq <- deq %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, 
             color = disturbance, fill = disturbance)) +
  geom_hline(yintercept = 0, col = 'red') +
  geom_point() +
  geom_area(alpha = 0.2) +
  xlab('Time (years)') +
  ylab('Distance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols) +
  scale_fill_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_Ieq <- Ieq %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = Ieq.structure, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Cummulative integral of\ndistance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
cowplot::plot_grid(g_agb, g_resilience, g_deq, g_Ieq, labels = LETTERS[1:4])
```

@@

The effect of biodiversity on ecosystem resilience can be split in complementarity and selection effect using @Loreau2001 partitioning:

@@

Biodiversity net effect $NE$ is based on the difference between ecosystem variable $X$ observed value $X_O$ within the community mixture of species and its expected value $X_E$ if species performance were equal to their performance in monocultures. This effect can be partitioned between complementarity effect $CE$, representing niche partitioning, positive interactions, and resource supply, and selective effect $SE$ due to dominant species pool driving the ecosystem. $N$ represents the total number of species, and $M$ the vector of monocultures performance. Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:

@@

$X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$. $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance. Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives). Whereas $SE$ become positive when dominant species outperform themselves in mixture than in monoculture, and negative when less dominant species outperform themselves in mixture than in monoculture [@Tobner2016]. But similarly, to resilience measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance. So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by undisturbed control ecosystem net effect $NE_C$ to measure treatments net effect resilience $R(NE_T)$:

@@

Resilience trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$. In order to do that, we simulated the 163 species individually representing 652 simulations of monoculture.

@@

Species richness was not directly link to $Ieq_{600}$. Increased functional diversity reduced cumulative integral from ecosystem distance to equilibrium after 600 years ($Ieq_{600}$). In addition, functional evenness was complementary reducing $Ieq_{600}$. Finally, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$. We found similar results for all disturbance levels and forest functioning (see [Appendix 5: Disturbance simulations]).

```{r gIeq, echo=FALSE, fig.cap='Ecosystem resilience after 600 years with taxonomic and functional diversity for different levels of disturbance. Cummulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008]. Solid lines and grey areas show fitted relationships with their mean and variance of the form $Ieq_600 = FDiv + \\epsilon$.', cache=TRUE, fig.width=12}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>% 
  ggplot(aes(x = FDiv, y = Ieq.structure, label = sim)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(shape = richness, color = FEve)) +
  facet_wrap(~disturbance, scales = "free", labeller = 'label_both') + 
  xlab('Functional diversity (FDiv)') +
  ylab('Cummulative integral from ecosystem 
       distance to forest structure equilibrium after 600 years') +
  scale_shape_discrete('Species\nrichness\n(SR)') +
  scale_color_continuous('Functional\nevenness\n(FEve)')
```

@@

## Complementarity and selection insure forest resilience

>>>>

---
title: Functional diversity improves tropical forest recovery from disturbance shifting from complementarity to selection effect (to rework)
author: "Sylvain Schmitt, Isabelle Maréchaux, Jerome Chave, Fabian Fischer, Camille Piponiot, Stéphane Traissac & Bruno Hérault"
date: '`r Sys.Date()`'
output:
  bookdown::html_document2:
    number_sections: false
    toc: true
    toc_float: yes
  bookdown::word_document2: default
  github_document: default
  bookdown::pdf_document2:
    includes:
      before_body: ./template/doc_prefix.tex
      in_header: ./template/preamble.tex
    number_sections: false
    toc: false
    keep_tex: true
linestretch: 1.5
csl: /home/sylvain/Documents/Bibliography/csl/mee.csl
bibliography: /home/sylvain/Documents/Bibliography/library.bib
link-citations: yes
---

```{r config, message=FALSE, warning=FALSE, include=FALSE}
rm(list = ls()); invisible(gc())
library(knitr)
library(kableExtra)
if(knitr:::is_html_output()) options(knitr.table.format = "html") 
if(knitr:::is_latex_output()) options(knitr.table.format = "latex") 
library(tidyverse)
library(broom)
theme_set(bayesplot::theme_default())
opts_chunk$set( echo = F, message = F, warning = F, 
                fig.height = 8, fig.width = 8,
                cache = T, cache.lazy = F)
```

@@

Natural disturbances caused by tree death play a major role in tropical forest functioning through gap-phase regeneration [@Veblen1992]. Climate change and direct human activities are currently deeply modifying forest disturbance regimes [@Davidson2012; @Lewis2015]. If deforestation has been rightly the focus of worldwide attention, forest degradation has been less studied and quantified through tropics [@Putz2010; @Simula2009]. However, degraded forests cover areas over ten times larger than deforestation [@Herold2011], a majority of which is being logged for timber production [@Blaser2011]. Forest degradation can trigger severe biodiversity loss [@Burivalova2014] and is estimated to account for 20.7 billion of tons of carbon dioxide emissions per year [@Pearson2017], representing 20% of greenhouse gas emissions in the Brazilian Amazon [@Asner2005]. Understanding the drivers of tropical forest resilience to disturbance is thus critical.

Resilience concept has met numerous definitions, but resilience in ecology can encompass both ecosystem ability to adsorb and recover from disturbance, resistance and recovery respectively [@Gunderson2010]. Both exogenous (e.g. climate, disturbance features) and endogenous (e.g. forest structure and composition) factors can drive ecosystem recovery after disturbance. Among the former, logging intensity has been reported to strongly control and increase time-to-recovery of post-logging tree carbon, subsequent animal biodiversity loss, and change in tree species composition [@Burivalova2014; @DeAvila2015; @Rutishauser2015]. Among endogenous drivers, forest pre-disturbance composition has been identified as a key driver of forest recovery [@Herault2018], as different species respond differently to perturbations depending on their ecological strategies and functional properties [@herault_functional_2011]. Beyond species identity, tree community diversity may also foster forest resilience to perturbations [@Cadotte2011; @Ives2007; @Loreau2013].

Two types of mechanisms can underlie an effect of biodiversity on ecosystem properties: complementarity and selection effects [@Loreau2001]. Complementarity effect results from resources partitioning and niche differentiation, and facilitation, leading to a more efficient use of available resources. Selection effect results from a higher probability of more diverse community to include particularly efficient species regarding ecosystem process or stability. Moreover the sampling effect additionally implies redundancy and a lower risk to lose important functional traits, further improving resilience of species rich-forest [@Hooper2005a; @Tilman2001].

The effect of biodiversity on ecosystem functioning has been intensely studied over the last decades, through theoretical work [@Loreau2010], experiments [@Cardinale2009], or observations along natural gradient of biodiversity [@Chisholm2013; @Liang2016; @Paquette2011; @Poorter2017]. However, experiments have mostly been conducted on grasslands [@Hector2005], while studies relying on forest stand inventories often lack repeatability [@Schnitzer2016], rarely include logged forests [@Sist2015], and monitoring temporal coverage remains too low to assess long-term effect as trees are long-lived organisms [@Herault2018]. Forest simulators have thus proven useful to investigate the effect of biodiversity on forest ecosystem functioning through virtual biodiversity experiments [@Morin2011; @Morin2014; @Sakschewski2016], but the problem of biodiversity representation beyond a few plant functional types in models of tropical forests have long been an impediment to apply this approach to this species-rich ecosystem [@Kazmierczak2014; @Kohler1998; @Purves2008].

In the present study, we take advantage of recent developments made in the forest simulator TROLL [@Chave1999], an individual-based and spatially-explicit model of tropical forest, that allow to deal with high diversity levels through a parameterization using plant functional traits [@Marechaux2017]. Consequently, TROLL offers the opportunity to study biodiversity-ecosystem functioning *in silico* in tropical forest communities.

We used a simulation approach to assess the effect of species richness and functional composition on long-term forest recovery from disturbances of different levels. We manipulated species number and identity across a large number of simulations, and measured the biodiversity net effect on simulated forest recovery after disturbance. In order to further investigate the mechanisms underlying such potential effect of biodiversity, we partitioned it into complementarity and selection effects. Based on the general hypothesis of a positive relationships between biodiversity and productivity [@Liang2016], we hypothesized that more diverse forest will recover quicker to a disturbance due to an increased productivity.

@@

TROLL is an individual-based and spatially explicit forest dynamic model in which the life cycle of individual trees (recruitment, growth, seed production and death) $>1cm$ diameter is simulated. Trees grows in a forest light environment explicitly computed within a $1m^3$ resolution voxel grid. Seedlings below the $1cm$ size class are not modeled explicitly, but are part of a seed/seedlings pool. A maximum of one tree can establish in each $1*1m$ pixel. Each tree is defined by its age, diameter at breast height (dbh), height (h), crown radius (CR), crown depth (CD) and leaf area (LA). Tree geometry is calculated with allometric relationships. Additionally, each tree is flagged with a species label inherited from the parent tree through the seedling recruitment. To each species label is associated a number of species-specific traits (Table \@ref(tab:traits)) from which processes are simulated using up-to-date established relationships drawn from the literature. The model thus represents tree biodiversity in a realistic way, where data requirement and model outputs parallel current trait collection efforts and field inventories, respectively.

Carbon assimilation is computed over half-hourly period of a representative day per month using the Farquhar, von Caemmerer, and Berry model of photosynthesis [@Farquhar1980].  Photosynthetic capacities, as well as leaf respiration rate, are computed from species-specific leaf concentration of nitrogen (N) and phosphorous (P), and leaf mass per area [LMA; @Atkin2015; @Domingues2010]. The net amount of assimilated carbon available for growth is then partitioned among the different plant organs using empirical factors, and converted into increments of stem volume and leaf area using species-specific wood-density and LMA respectively. Leaf demography is simulated using species-specific leaf lifespan derived from LMA. Following an underestimation of leaf lifespan for low LMA species, we designed our own leaf lifespan allometry, specific to tropical tree species, to be used in subsequent analysis (see supplementary materials S1). Natural tree deaths are stochastic events whose background rate is negatively correlated with wood density, as observed pan-tropically [@kraft_functional_2010; @poorter_are_2008; @Wright2010]. Additionally, tree can die due to carbon starvation or treefalls. Below-ground processes, herbaceous plants, epiphytes and lianas are not simulated in this version of TROLL. The source code is written in C++ and available at https://github.com/TROLL-code/TROLL. A further detailed description of the model is available in @Marechaux2017.

@@

A new disturbance module was developed for this study. At a given iteration $disturb_{iter}$, individual trees are randomly picking following a uniform law and without any individuals nor species targeting. Selected individuals are then removed without triggering a treefall to avoid any side effect. The operation is repeated until the disturbance result in a defined removed basal area ($disturb_{intensity}$).

## Simulation experiment

In order to assess the role of biodiversity on ecosystem recovery from disturbance, we  simulated forest trajectories for different levels of disturbance intensity and tree community compositions. For each simulation, we first simulated forest regeneration from bare soil for 600 years [pre-disturbance step, @Marechaux2017], then simulated the disturbance described as above ($disturb_{iter}$), and finally let the simulated forest community freely recover from the disturbance during 600 years (post-disturbance step). Disturbance intensity was quantified by the percentage of basal area loss, here fixed at 0% (control), 25%, 50% and 75%. For each level of disturbance, we performed simulations differing in the original number of species ($SR=15,25,125$) and composition. For each level of species richness, we created 1000 virtual communities by randomly picking desired number of species among the regional species pool. We used 20 communities for simulations with a large range of convex hull volume $CVH$ but with a community weighted mean close to the regional species pool community weighted mean. This led to 240 simulations varying in community composition and disturbance intensity ($3~SR*20~CHV*4~ levels~of~disturbance$).

All simulations were made with a parameterization for an Amazonian forest of French Guiana, with 163 parameterized species, as in @Marechaux2017.

## Analysis

### Quantifying simulated community diversity

We focused on functional richness with convex hull volume $CHV$ and functional mean with community weighted mean $CWM$ to define simulation communities. Whereas species and functional diversities were assessed for each simulation at the end of the pre-disturbance step with species richness ($SR$) and @villeger_dynamique_2008 indices ($FRIC$, $FEve$, and $FDiv$).

### Ecosystem recovery from disturbance

**We described the simulated tropical forests from changes in values of forest structure (aboveground biomass ($AGB$ in $ton~C.ha^{-1}$), basal area ($BA$ in $m^2.ha^{-1}$), total number of stem ($N$), number of stem above $10~cm$ diameter ($N10$), and number of stem above $30~cm$ diameter ($N30$)) and in values of forest functioning (growth primary productivity ($GPP$ in $MgC.ha^{-1}$), net primary productivity ($NPP$ in $MgC.ha^{-1}$), tree autotrophic respiration in day ($Rday$ in $MgC.ha^{-1}$) and tree autotrophic respiration in night ($Rnight$ in $MgC.ha^{-1}$)).**

The recovery of the simulated forest ecosystem at a time $t(R(t))$ regarding a given properties $X$ was defined ad follows:

\begin{equation}
  R\left(t\right)= \frac{X_T(t)}{X_C(t)}
  (\#eq:Recovery)
\end{equation}

where $X_T(t)$ and $X_C(t)$ are the ecosystem metric values at time $t$ in the disturbed simulation ($disturb_{intensity} = 25\%, 50\%, 75\%$) and in the undisturbed control simulation ($disturb_{intensity} = 0\%$). Note that demographic stochasticity in the model induces a low variability among replicated simulations with the same set-up so that **XX** [@Marechaux2017]. When $R(t)$ approaches $R_{eq}=1$, the disturbed and control systems are considered equivalent, indicating a perfect recovery. We then calculated the euclidean distance to this state of perfect recovery as $d_{eq}(t) = \sqrt{(R_{eq}-R(t))^2}$. **Ecosystem euclidean distance to equilibrium distance was calculated in a multi-dimensional space for (the two functions described above: (i) forest structure (AGB, BA, N, N10, and N30) and (ii) forest functioning (GPP, NPP, Rday, and Rnight). We then integrate the euclidean distances over time to get our ecosystem recovery indicators.** (Figure \@ref(fig:datatrans)). We used the cumulative integral after 600 years $Ieq_{600}$ as a measurement of ecosystem recovery. A high value of $Ieq_{600}$ indicating a slow recovery. The influence of pre-disturbance species richness ($SR$), functional evenness ($FEve$) and functional diversity ($FDiv$) on ecosystem recovery index $Ieq_{600}$  was finally investigated by linear regression ($Ieq_{600} \sim SR + FEve + FDiv$).

```{r datatrans, echo=FALSE, fig.cap='Ecosystem outputs data transformation. Ecosystem outputs (**A**) are normalized by the control value over time to calculate recovery (**B**); recovery of different ecosystem outputs is then used in a multidimensional space to caluclate ecosystem distance to equilibrium (**C**); finally distance at equilibrium is integrated over time in a cumulative sum (**D**). ', cache=TRUE, fig.height=8, fig.width=12, message=FALSE}
cols <- RColorBrewer::brewer.pal(3, "YlOrRd")
g_agb <- ecosystem %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb/1000, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Aboveground biomass (tonC/ha)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                       values=c("#999999", cols))
g_resilience <- resilience %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = agb, color = disturbance)) +
  geom_hline(yintercept = 1, col = 'red') +
  geom_point() +
  xlab('Time (years)') +
  ylab('Resilience of aboveground biomass') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_deq <- deq %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = structure, 
             color = disturbance, fill = disturbance)) +
  geom_hline(yintercept = 0, col = 'red') +
  geom_point() +
  geom_area(alpha = 0.2) +
  xlab('Time (years)') +
  ylab('Distance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols) +
  scale_fill_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
g_Ieq <- Ieq %>% 
  left_join(., doe, by = 'sim') %>%
  filter(richness == 125) %>% 
  filter(cvh == 1) %>% 
  ggplot(aes(x = time, y = Ieq.structure, color = disturbance)) +
  geom_point() +
  xlab('Time (years)') +
  ylab('Cummulative integral of\ndistance to forest structure equilibrium\n(AGB, BA, N, N10 and N30)') +
  scale_color_manual('Disturbance\nintensity\n(% of BA)',
                     values = cols)
cowplot::plot_grid(g_agb, g_resilience, g_deq, g_Ieq, labels = LETTERS[1:4])
```

@@

The net effect of biodiversity on ecosystem recovery (NE) is defined as the difference between the ecosystem recovery of the mixed-species community and the one expected under the null hypothesis that there is no effect of biodiversity, i.e. ecosystem recovery for the mixed-species community equals the abundance-weighted sum of that in monoculture [@Loreau2001]. We then split this net biodiversity effect in complementarity and selection effect using @Loreau2001 partitioning:

@@

where $N$ represents the total number of species, and $M$ the vector of monoculture performance. Both metrics depend on the variation of relative ecosystem variable $\Delta RX$:

@@

where $X_{sp}$ is the ecosystem variable value for one species either in mixture $X_{sp}(mixture)$ or in monoculture $X_{sp}(monoculture)$. $P_{sp}$ is the proportion of the species in the mixture represented by species relative abundance. Consequently, $CE$ averages diversity effects of all species presents in the mixture (both negatives and positives). Whereas $SE$ become positive when dominant species outperform themselves in mixture than in monoculture, and negative when less dominant species outperform themselves in mixture than in monoculture [@Tobner2016]. But similarly, to recovery measurement, biodiversity net effect $NE$ is a dynamic equilibrium and vary over time without disturbance. So in order to correctly assess selection and complementarity effect in answer to disturbance, we normalized it by ecosystem net effect $NE_C$ from the undisturbed control simulation ($disturb_{intensity} = 0\%$) to measure treatments net effect recovery $R(NE_T)$:

@@

Recovery trajectories of ecosystem variable after disturbance were partitioned between complementarity effect $CE$ and selection effect $SE$. In order to do that, we simulated the 163 species individually representing 652 simulations of monoculture.

@@

Species richness was not significantly influencing $Ieq_{600}$ with forest structure and had a small positive effect on forest production (Table \@ref(tab:tgIeq)). Increased functional diversity reduced cumulative integral from ecosystem distance to equilibrium after 600 years ($Ieq_{600}$) and functional evenness was complementary reducing $Ieq_{600}$ (Figure \@ref(fig:gIeq) and Table \@ref(tab:tgIeq)). Finally, low species richness could result in variant $Ieq_{600}$, but increased species richness resulted in increased functional diversity and consequently lower $Ieq_{600}$. We found similar results for all disturbance levels (see [Appendix 5: Disturbance simulations]).

```{r gIeq, echo=FALSE, fig.cap='Ecosystem recovery after 600 years with taxonomic and functional diversity for different levels of disturbance. Cumulative integral from ecosystem distance to forest structure equilibrium after 600 years was represented against functional diversity [FDiv, @villeger_new_2008] for different level of disturbance (25, 50 and 75% of total basal area); dot shapes represents the species richness whereas dot color represents functional evenness [FEve, @villeger_new_2008]. Solid lines and grey areas show fitted relationships with their mean and variance of the form $Ieq_600 = FDiv + \\epsilon$.', cache=TRUE, fig.width=12}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  filter(time == 600) %>% 
  ggplot(aes(x = FDiv, y = Ieq.structure, label = sim)) + 
  geom_smooth(method = 'lm', alpha = 0.3, colour= 'grey') +
  geom_point(aes(shape = richness, color = FEve)) +
  facet_wrap(~disturbance, scales = "free", labeller = 'label_both') + 
  xlab('Functional diversity (FDiv)') +
  ylab('Cummulative integral from ecosystem 
       distance to forest structure equilibrium after 600 years') +
  scale_shape_discrete('Species\nrichness\n(SR)') +
  scale_color_continuous('Functional\nevenness\n(FEve)')
```

```{r tgIeq, cache=T}
Ieq %>% 
  left_join(., doe, by = 'sim') %>% 
  mutate(SR = as.integer(richness)) %>% 
  select(disturbance, Ieq.structure, Ieq.production, SR, FEve, FDiv) %>% 
  reshape2::melt(id.vars = c("disturbance", "SR", "FEve", "FDiv"),
                 variable.name = "Ieq.type", value.name = "Ieq.value") %>% 
  group_by(Ieq.type, disturbance) %>%
  do(fit = lm(Ieq.value ~ SR + FEve + FDiv, data = .)) %>% 
  tidy(fit) %>% 
  ungroup() %>% 
  filter(term != "(Intercept)") %>% 
  mutate(`P-value` = ifelse(p.value < 0.001, "***",
                       ifelse(p.value < 0.01, "**",
                              ifelse(p.value < 0.05, "*",
                                     ifelse(p.value < 0.1, ".", "n.s."))))) %>% 
  mutate(Coefficient = round(estimate,2)) %>% 
  mutate(Recovery = gsub("Ieq.", "", Ieq.type)) %>% 
  rename(Variable = term, `Disturbance intensity` = disturbance) %>%
  select(Recovery, `Disturbance intensity`, Variable, Coefficient, `P-value`) %>%
  kable(caption = "Result of linear regression testing the effect of species richness and functional diversity indices on the recovery index by recovery type and disturbance intensity.") %>% 
  kable_styling(full_width = F) %>%
  column_spec(1, bold = T) %>%
  collapse_rows(columns = 1:2)
```

@@

## Complementarity and selection insure forest recovery
